Empowering the use of preprints in ecology and evolution: an exploration of the first 1000 *EcoEvoRxiv* preprints

### 2023-10-11 ms/ms.qmd
INtro editing as I didn't like a few things

Preprints attempt to break down barriers to scientific publishing by: 1) increasing the speed, visibility and output of research findings; 2) encouraging constructive and critical feedback from a research community; 3) helping establish precedence of research findings; and 4) removing financial barriers to publication [@Proulx2013 and others]. Importantly, *EcoEvoRxiv* is a community driven preprint server that has paved the way for new initiatives that attempt to further break down barriers to scientific dissemination and improve research practice, including accepting multilingual preprints (i.e., Spanish and Portuguese), recognising and accepting registered reports and promoting the use of peer review and community discussion to improve the quality of preprints and speed up their peer-reviewed publication (e.g., Peer Community In -- PCI).

Preprint servers, such as *EcoEvoRxiv*, can empower researchers to make their findings more accessible, open and transparent but only if they are used, and used well. Significant barriers to preprint use still exist globally – from worry about XXX to XXX. Nonetheless, we lack a systematic understanding of the factors that influence preprint use in ecology and evolution which can help inform future initiatives and help us begin to break down barriers to their use in our field. Here, we explore the first 1000 preprints uploaded to *EcoEvoRxiv* to understand preprint practices and use in our field, focusing on our attention on understanding: 1) how long it takes for a preprint to become published, and how many preprints remain unpublished; 2) what countries make use of preprints; 3) how career stage and gender impact preprint use; 4) whether data and code are more likely to be shared in preprints; and 5) the extent to which authors make use of preprint servers for registered reports and community driven peer review.

>>>>

Preprints attempt to break down barriers to scientific publishing by: 1) increasing the speed, visibility and output of research findings; 2) encouraging constructive and critical feedback from a research community; 3) helping establish precedence of research findings; and 4) removing financial barriers to publication [@Proulx2013 and others]. Importantly, *EcoEvoRxiv* is a community driven preprint server that has paved the way for new initiatives, including accepting multilingual preprints (i.e., Spanish and Portuguese), recognising and accepting registered reports and promoting the use of peer review and community discussion to improve the quality of preprints and speed up their peer-reviewed publication (e.g., Peer Community In -- PCI).

Preprint servers, such as *EcoEvoRxiv*, can empower researchers to make their findings more accessible, open and transparent but only if they are used well. Significant barriers to preprint use still exist globally – from worry about XXX to XXX. Nonetheless, we lack a systematic understanding of the factors that influence preprint use in ecology and evolution which can help inform future initiatives and allow us to begin to break down barriers to their use in our field.

Here, we explore the first 1000 preprints uploaded to *EcoEvoRxiv* to understand preprint practices and use in our field, focusing on our attention on understanding: 1) how long it takes for a preprint to become published, and how many preprints remain unpublished; 2) what countries make use of preprints; 3) how career stage and gender impact preprint use; 4) whether data and code are more likely to be shared in preprints; and 5) the extent to which authors make use of preprint servers for registered reports and community driven peer review.
